**Machine learning identifies unique taxa differentiating proximal and distal

### 2017-05-23 submission/manuscript.Rmd
manuscript updates

#### Microbial membership and diversity of the proximal and distal colon 
Proximal and distal lumenal and mucosal samples were collected from 20 healthy humans that had not undergone bowel preparation prior to the procedure (Figure 1). Participants also collected stool at home one week prior to the procedure. To characterize the bacterial communities present at these sites, 16S rRNA gene sequencing was performed on extracted DNA from each sample. The relative abundance of each phyla for each site is shown in Figure 2A. This data combines the samples from all participants in the study. Each site is primarily dominated by Firmicutes and Bacteriodetes, consistent with known variability in human microbiome research. ((Is anything statistically significant here? use Doherty's code)). Likewise, samples had varying levels of diversity at each site, irrespective of the individual (Figure 2B). That is, while some individuals had a more diverse right mucosa, some had a more diverse left mucosa, and so on. Thus there was no clear pattern of changes in microbial diversity along the gut axis.

@@

### Discussion
Here we identified bacterial taxa that are specific to the lumen and mucosa of the right and left human colon from samples collected during unprepared colonoscopy. We found that all locations contained a range of phyla and a range of diversity, but that there was a wide variability between subjects. Pairwise comparisons of each of the sites revealed that the right mucosa and lumen were most similar to each other. Further, comparison of colonoscopy-collected samples with samples collected from stool at home showed that the left lumen is most similar to stool at exit. Random Forest algorithms built on OTU abundances from each sample identified microbes that are particular to each location of the colon.

@@

discuss host response papers

>>>>

#### Microbial membership and diversity of the proximal and distal colon 
Proximal and distal lumenal and mucosal samples were collected from 20 healthy humans that had not undergone bowel preparation prior to the procedure (Figure 1). Participants also collected stool at home one week prior to the procedure. To characterize the bacterial communities present at these sites, 16S rRNA gene sequencing was performed on extracted DNA from each sample. The relative abundance of each phyla for each site is shown in Figure 2A. This data combines the samples from all participants in the study. Each site is primarily dominated by _Firmicutes_ and _Bacteriodetes_, consistent with known variability in human microbiome research. ((Is anything statistically significant here? use Doherty's code)). Likewise, samples had varying levels of diversity at each site, irrespective of the individual (Figure 2B). That is, while some individuals had a more diverse right mucosa, some had a more diverse left mucosa, and so on. Thus there was no clear pattern of changes in microbial diversity along the gut axis.

@@

### Discussion
Here we identified bacterial taxa that are specific to the lumen and mucosa of the right and left human colon from samples collected during unprepared colonoscopy. We found that all locations contained a range of phyla and a range of diversity, but that there was a wide variability between subjects. Pairwise comparisons of each of the sites revealed that the right mucosa and lumen were most similar to each other. Further, comparison of colonoscopy-collected samples with samples collected from stool at home showed that the left lumen is most similar to stool at exit. Random Forest algorithms built on OTU abundances from each sample identified microbes that are particular to each location of the colon. Finally, we were able to detect some bacterial OTUs associated with colorectal cancer in our healthy patient cohort. Using unprepped colonoscopy and machine learning, we have identified bacterial phyla specific to the healthy proximal and distal human colon.

We identified a range of Simpson diversity values between patients and between each anatomical site. As this measurement takes both number of species and abundance of species into account, it seems that there is a varied level of each component along the human colon. There was no clear pattern of diversity when comparing patients- some patients had a higher level of diversity in the proximal than the distal, some patients had the opposite. An early study of microbial diversity and membership of three human subjects found a varied level of diversity in each subject, but that subjects were overall more similar to themselves than to each other (Eckberg 2005). When ordinating the subject communities to compare the right and left sides, we found that more than half of the samples from the same patient clustered near each other (Figure 2C). However, there were a few patients whose samples were very different from each other and did not ordinate in that manner. One explanation for this ____ - human diversity, sampling locations arent perfectly matched (but so what?)

@@
